---
title: "Getting and Cleaning finalData Course Project"
output: 
      html_document:
          keep_md: true
      md_document:
          keep_md: true
---

 

 1. Merges the training and the test sets to create one data set.
 create a new directory if not exist
```{r}
if(!file.exists("./projectdata")){
      dir.create("./projectdata")
}
```

 Download and unsiz the file
```{r}
dataUrlLing <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(dataUrlLing, destfile="./projectdata/FileData.zip", method="curl")
unzip(zipfile="./projectdata/FileData.zip",exdir="./projectdata")
```

 In this section, we will  merge the following files:
 1. test/subject_test.txt
 2. train/subject_train.txt
 3. test/X_test.txt
 4. train/X_train.txt
 5. test/y_test.txt
 6. train/y_train.txt

"|-------------------------------------------------------------------"
"|            subject         |     activity       |   feature.txt  |"  
"|----------------------------|--------------------|----------------|"
"|                            |                    |                |"
"|       subject_test.txt     |    y_test.txt      |   X_test.txt   |"
"|                            |                    |                |
"|----------------------------|--------------------|----------------|
"|                            |                    |                |
"|      subject_test.txt      |   y_train.txt      |   X_test.txt   | 
"|                            |                    |                |
"| ------------------------------------------------------------------

........................................................................
|            subject         |     activity       |   feature.txt      |  
........................................................................
|     subject_test.txt       |    y_test.txt      |   X_test.txt       |
........................................................................
|      subject_test.txt      |   y_train.txt      |   X_test.txt       | 
........................................................................

 This fnction will return the data ride from the file.
```{r}
library(data.table)
library(plyr)
readFileInTable <- function( subDirectory="", fileName="") {
      filePath <- file.path("./projectdata" , "UCI HAR Dataset")
      read.table(file.path(filePath, subDirectory , fileName ),header = FALSE)
}
```

Read the header column names from activity_labels.txt" and features.txt
```{r}
activityLabelsData  <- readFileInTable("", "activity_labels.txt")
featuresLabelsData  <- readFileInTable("" , "features.txt")
```


We will first read all the files into different data.table variable as follow:
```{r}
featuresTestData  <- readFileInTable("test" , "X_test.txt" )
featuresTrainData <- readFileInTable("train", "X_train.txt")

activityTestData  <- readFileInTable("test" , "Y_test.txt" )
activityTrainData <- readFileInTable("train", "Y_train.txt")

subjectTrainData  <- readFileInTable("train", "subject_train.txt")
subjectTestData   <- readFileInTable("test" , "subject_test.txt")
```

Merge the files
```{r}
features <- rbind(featuresTestData, featuresTrainData)
activity <- rbind(activityTestData, activityTrainData)
subject  <- rbind(subjectTestData,  subjectTrainData)

names(features) <- featuresLabelsData$V2
names(activity) <- c( "activity")
names(subject) <- c( "subject") 

mergeSubActionData <- cbind(subject,  activity)
finalData          <- cbind(features, mergeSubActionData)
```

create a feature names logical vector 
```{r}
filterNameData <- sapply(featuresLabelsData[,2], function(s) { grepl('(mean|std)', s, ignore.case = T) })
finalData <- finalData[, filterNameData]
```

2. Extracts only the measurements on the mean and standard deviation for each measurement. 
```{r}
subFeaturesData<-featuresLabelsData$V2[grep("mean\\(\\)|std\\(\\)", featuresLabelsData$V2)]
selectedNames<-c(as.character(subFeaturesData), "subject", "activity" )
finalData<-subset(finalData,select=selectedNames)
str(finalData)
```
 3. Uses descriptive activity names to name the activities in the data set
```{r}
finalData$activity[finalData$activity==1] <- "Walking"
finalData$activity[finalData$activity==2] <- "Walking Upstairs"
finalData$activity[finalData$activity==3] <- "Walking Downstairs"
finalData$activity[finalData$activity==4] <- "Sitting"
finalData$activity[finalData$activity==5] <- "Standing"
finalData$activity[finalData$activity==6] <- "Laying"
```

4. Appropriately labels the data set with descriptive variable names. 
In this section we will rename all the variable that does not have a proper name using regex
```{r}
names(finalData) <- gsub("^t",       "time",          names(finalData))
names(finalData) <- gsub("^f",       "frequency",     names(finalData))
names(finalData) <- gsub("BodyBody", "Body",          names(finalData))
names(finalData) <- gsub("Gyro",     "Gyroscope",     names(finalData))
names(finalData) <- gsub("Acc",      "Accelerometer", names(finalData))
names(finalData) <- gsub("Mag",      "Magnitude",     names(finalData))
```

5. From the data set in step 4, creates a second, independent tidy data set 
```{r}
newAggregateData <- ddply(finalData, .(activity, subject), numcolwise(mean))
write.table(newAggregateData, file = "tidydatafile.txt",row.name=FALSE)
```
